Is anchoring better?

### 2022-05-25 manuscript/body.tex
Autocommit
Working on BMB's feedback

@@

In many applications which use simple or generalized linear (mixed) models, outcome predictions or probabilities, are often of interest. Predictions provide a great way to summarize what the regression model is telling us and very useful for interpreting and visualizing model estimates. For example, in logistic regression models, the coefficient estimates are usually not easy to interpret and less informative. Since logistic models are nonlinear (due to the nonlinear link function), in multivariate models with interactions, the magnitude of the effect of change in the outcome depends on the values of the predictor of interest and other predictors. Hence, the conclusions one can make about the estimated effects greatly depends on how well we choose the values of the other predictors, i.e., \emph{non-focal} predictors.

Both simple and generalized linear (mixed) models (GL(M)Ms) can examine very complex relationships, including nonlinear relationships between response and predictors, interactions between predictors (and via splines for example), and nonlinear transformations via link functions due to their flexibility. This flexibility comes at a cost, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest. Also, coefficient estimates of models involving nonlinear link functions or interactions lose their direct interpretation \citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore the \emph{effects} of predictors on the predictions or probabilities of the outcome at various level of the predictor of interest -- \emph{prediction and effects} \citep{fox2009effect, leeper2017package, lenth2018package}. For example, as a way to plan, public health officials may want to know the effects of household income, wealth index, etc, on the predicted probability of having improved water services among slum dwellers.

When visually presented, prediction and effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the predictors whose estimates are usually, but not always, a subject to less clarity of interpretation. Further, generating predictor and effect plots together with the associated confidence intervals for regression models has a number of challenges. In particular:
\begin{enumerate}
\item choice of representative values of \emph{focal} predictor and the \emph{reference point} for non-focal predictors especially in multivariate models
\item  propagation of uncertainty -- appropriate choice of \emph{anchor} for computing confidence intervals; and how to incorporate the uncertainty due to non-focal predictors 
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)Ms)
\end{enumerate}

The most common way of dealing with the first challenge is taking unique levels of the focal (predictor of interest) predictor if discrete or taking appropriately sized quantiles (or bins) if continuous, and then calculating the predictions while holding non-focal (other predictors other than the focal predictor) predictors at their reference point (e.g., means) \citep{hanmer2013behind}. This generates -- \emph{predictions}, \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to this quantity as predictions since it should, for example, tell us what we would expect the response to be at a particular value or level of the predictor, for an ``average case''. More specifically, we compute the expected outcome by meaningfully holding non-focal predictors constant (or averaged in some meaningful way) while varying the focal predictor, with the goal that the outcome expected prediction represents how the model responds to the changes in the focal predictor. On the other hand, effects differs with predictions in the sense that the former is about how we describe the uncertainty around the latter and depends on the anchor.

The commonly used \proglang{R} software packages (\pkg{emmeans} and \pkg{effects}) for generating predictions, by default, use the average of non-focal predictors as the reference point. However, there are a number of choices one can make when considering this approach -- for example, in the presence of interaction, averaging the interactions (averaging product of interacting model variables) versus product of the averages of the interacting model variables (default for \pkg{emmeans} and \pkg{effects}). We claim that neither of these two approaches is the most appropriate but we show that averaging the interaction closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal predictor:

%
\begin{align}
y &= \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3x_3 + \epsilon \label{eq:simple_inter_higher_no_interaction}\\
y &= \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3x_3 + \beta_{23}x_2x_3 + \epsilon \label{eq:simple_inter_higher}
\end{align}
%

We first simulate data with the two models, such that $x_{1,2,3} \sim \mathrm{Normal}(0, 1)$, $\beta_0 = 5$, $\beta_1 = -3$, $\beta_2 = 1$, $\beta_3 = 2$ and $\beta_{23} = 5$, and then compare the predictions from the \pkg{emmeans}, \pkg{effects} and our proposed alternative (\pkg{varpred}) to the ``true'' predictions and observed average i.e., $\bar{y}$, as shown in Fig~\ref{fig:justify_plots}.

\begin{figure}[!h]
\centering
\includegraphics[width=0.9\textwidth]{justify_preds.inter.pdf}
\caption{{\bf A comparison of \pkg{emmeans}, \pkg{effects} and \pkg{varpred} predictions for $x_1$ on $y$ for models with and without interactions.}
The horizontal blue, green and black lines are the mean predictions, i.e., $\bar{\hat{y}}$, the red ones are the means of the simulated $y$, i.e., $\bar{y}$, while the vertical dotted grey lines are the means of the corresponding focal predictors. The grey points are the binned simulated $y$. The trend lines represents the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other predictors at their average for the three mthods and the ``true'' predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches closely matches the ``true'' predictions. B: Even with the simple interaction between non-focal predictors, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and ``true'' predictions, respectively, in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not, the proposed \pkg{varpred}.}
\label{fig:justify_plots}
\end{figure}

@@

Sometimes, one may be interested in the uncertainties associated with the focal predictor of interest only, excluding other uncertainties due to other non-focal predictors -- effects. However, currently, the two packages do not provide straightforward way do achieve this. To illustrate this, we generate predictions of $x_2$ from model~\ref{eq:simple_inter_higher_no_interaction} together with the associated $95\%$ confidence bands, as shown in Fig~\ref{fig:justify_ci_plots}.

\begin{figure}[!h]
\centering
\includegraphics[width=0.9\textwidth]{justify_preds.isolate.pdf}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal predictor, i.e., the model center, corresponds to the effects of the focal predictor from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal predictors. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal predictor.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence bands are much wider because they include uncertainties associated with the intercept and non-focal predictors, but narrower and crosses at the mean of the focal predictor, i.e., \emph{center point or mean-based}, in \pkg{varpred}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal predictor we are more certain about, i.e., mean of the model variable corresponding to the focal predictor (center point or mean-based). For simple models, the point where the confidence bands crosses is the model center and it corresponds to anchor we choose, in this case, the mean of the focal predictor.

When dealing with nonlinear link functions or models with multiple non-focal predictors, the correct predictions, for example, are even much harder to estimate. One approach is to make predictions on the transformed scale (linear predictor scale), and then back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation. In particular, bias in expected mean prediction induced by nonlinear transformation of the response variable can lead inaccurate predictions. An alternative to mean-based reference point, i.e., averaging non-focal predictors, is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictors and then averaging across the values of the focal predictor \citep{hanmer2013behind}.

@@

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{justify_anchors.combined.pdf}
\caption{{\bf A comparison of min-based and center- (mean-) based anchors.} The center-based effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

We would like to summarize by talking about predictions vs. effects. Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor.

The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only -- thus, for example, if an effect is barely significant (P exactly 0.05), one end of the CI will be exactly a horizontal line for the effects plot, but not for the prediction plot. Also, the uncertainty curves for the effect, but not for the prediction, depend on the ``anchor''.

@@

To get an intuition of how conditioning on the mean values of non-focal predictors work, suppose we are interested in predictions at the levels of a particular predictor, i.e., focal, $x_f$, from the set of predictors. As a starting point, assume that the model has no interaction terms. The idea is to choose a reference point for the values of non-focal predictors. For example, fixing the values of non-focal predictor(s) at some typical values -- typically determined by averaging (for now) in some meaningful way, for example, arithmetic mean  for continuous and average over the levels of the factors for categorical non-focal predictors. We refer to this as centered or mean-based approach. The most convenient way to achieve this is by constructing $\bX^\star$ by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.

Consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function (in the case of simple linear model), where $\boldmu$ is the expected value of response variable $y$. Let $\hat{\bbeta}$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\boldsymbol{\Sigma} = V(\hat{\bbeta})$ of $\hat{\bbeta}$. Let $\mathbf{X^*}$ be the model matrix, inheriting most of its key properties, for example transformations on predictors and interactions from the model matrix, $\mathbf{X}$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\hat{\bbeta}$ is the prediction with respect to the focal predictor in question \citep{fox2009effect}.

An alternative formulation of $\hat{\boldeta}^\star$ involves expressing the linear predictor as the sum of the focal and non-focal predictors' linear predictor. In particular,

\begin{align}\label{eq:eta_mean}
\eta^\star(x_f, \nset{{\bar{x}^\star}}) &= \beta_f x_f + \sum \nset{\beta} \nset{{\bar{x}^\star}} \\
\hat{y}_f  &= g^{-1} \left(\eta^\star(x_f, \nset{{\bar{x}^\star}})\right)
\end{align}
where $\nset{{\bar{x}^\star}}$ are the appropriately averaged entries of non-focal predictors and $x_f$ is a vector of values of the focal predictors for a particular observation.

@@

Higher order terms such as interactions, splines, polynomials, etc., can be within the focal predictor, within (between) non-focal predictor(s), between the focal and non-focal predictor(s). To distinguish the three, suppose the model which describes the hypothetical simulated household size simulation based on a number of socio-demographic factors is

@@

In the first case, with $Age$ as the focal predictor, the interaction is within its polynomial terms (and 3 model variables associated with it). In this case, each of the polynomial terms (model variables) of the focal predictor are evaluated independently across the chosen levels of the focal predictor, $Age_i$. Specifically, the higher order terms are treated as additional columns of the model matrix evaluated with the same values of the focal predictor, while non-focal predictors are fixed at their reference point as discussed in the previous section. In this case, the predictions associated with $Age$ on the linear predictor scale becomes

@@

In the second case, with $Wealthindex$ as the focal predictor, the higher order interactions in non-focal predictor, $Age$, are treated simply as additional columns in the variable space (of model matrix) and appropriate choice of reference point applies just like in the models without interactions. For instance, in mean-based approach, we simply average all non-focal interaction terms. Thus

\begin{align*}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i
\end{align*}

Lastly, consider model~\ref{eq:simple_inter_higher}, with the interactions between the focal predictor, $x_2$, and non-focal predictor, $x_3$. In this case, the non-interacting predictors are treated as non-focal predictors, with appropriate reference points. However, for the interacting predictors, we first choose representative values for the focal predictor and some particular values of the interacting predictor similar to the way we choose values for the focal predictor or predetermined. In our example, suppose we pick $i$ and $j$ unique values of the focal predictor, $x_2$ and interacting predictor $x_3$, respectively, then the prediction on linear predictor scale is given by

@@

What about the confidence intervals (CI)? The limits of the confidence intervals are points, not mean values. In principle, every value of focal predictor has a different CI. The conventional way to compute variances for predictions is $\sigma^2 = \textrm{Diag}(\bX^\star \boldsymbol{\Sigma} \bX^{\star\top})$ \citep{lenth2018package, fox2009effect}, so that the confidence intervals are $\eta \pm q\sigma$, where $q$ is an appropriate quantile of Normal or t distribution. This generates predictions-styled CIs and incorporates all the uncertainties -- including the uncertainty due to the intercept and non-focal predictors.  But what if we are only interested in the uncertainty as a result of the focal predictor, so that the confidence intervals are $\eta \pm q \sigma_f$? We call this anchored CIs and generates effects-styled CIs.

Currently, commonly used \proglang{R} packages for constructing predictions, by default, do not exclude the uncertainties resulting from non-focal predictors when computing the CIs. A non-trivial way to exclude uncertainties associated with non-focal predictors in some of these packages is to provide a user defined variance-covariance matrix with the covariances of non-focal terms set to $0$ -- \emph{zeroing-out} variance-covariance matrix. This only works when the input predictors are \emph{centered} prior to model fitting, in case of numerical predictors, and even much complicated when the predictors are categorical. We first describe the variance-covariance based approach and then discuss our proposed method which is based on the model center and does not require input predictors to be scaled prior to model fitting.

\subsection*{Variance-covariance}

The computation of $\hat{\eta}^\star$ remains the same as described above. However, to compute $\sigma$, $\boldsymbol{\Sigma}$ is modified by \emph{zeroing-out} (the variance-covariance of all non-focal predictors are set to zero) variances of non-focal terms. Although this is the simplest approach, it requires centering of continuous, i.e., $x_c = x - \bar{x}$, predictors prior to model fitting and proper way to average categorical predictors.

@@

Let $\bxo$ be a centered model matrix containing appropriately chosen values of focal predictors and centered non-focal predictors, and let $\ba$ be the anchor matrix, with same dimensions as $\bxo$ and values in each columns repeated $n$ times. Also, let $\baf$ be the column of $\ba$ corresponding to the focal predictor, the rest of the columns corresponds non-focal predictors with same values as in $\bxo$. Any appropriate values (within the range of focal values) can be chosen for $\baf$ but for model center (center-anchored), we have $\bafc$ . Thus $\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top)$. We can see that $\forall ~\bxof=\baf$, $\sigma_f^2 = \boldsymbol{0}$. This is true for the center-anchored provided $\bxof=\bafc$. In other words non-focal terms (model variables) in $\bX^{\star}_{c} = (\bxo - \bafc)$ are all zero in simple models without interactions but are isolated (narrower) around the model center in models involving complex interactions. In addition, the computation of $\bX^{\star}_c$ impacts only on the intercepts and non-focal model variables, i.e., the slopes and variance of the focal predictors are not affected. This means that we can still generate effects without necessarily centering the predictors prior to model fitting.

@@

In many applications, it usually important to report the estimates that reflect the expected values of the untransformed response. This is not a major issue in simple linear models. However, when dealing with nonlinear link functions, it is even harder to generate correct predictions that reflect this due to the bias in the expected mean induced by the nonlinear transformation of the response variable. In such cases, bias correction is needed when back-transforming the predictions to the original scales. Most common approach for bias-adjustment is second-order Taylor approximation \citep{lenth2018package, duursma2003bias}. Another potential source of bias comes from additional non-focal predictors, especially in models with nonlinear link functions. Here, we describe and implement a different approach, whole-sample-based approach for bias correction.

@@

The most precise (although not necessarily accurate!) way to predict is to condition on values of the focal predictor and make predictions for all observations (members of the population) \citep{hanmer2013behind}. A key point is that the nonlinear transformation involved in these computations is always \emph{one-dimensional}; all of the multivariate computations required are at the stage of collapsing the multidimensional set of predictors for some subset of the population to a one-dimensional distribution of $\eta^\star(x_f, \nset{x})$, which is a function of focal predictor and the observed values of non-focal predictors, as opposed to the definition in Equation~\ref{eq:eta_mean}. More specifically:

>>>>

%% Main tex

@@

In applications that use regression models, outcome predictions are often of interest. Predictions provide a great way to summarize what the regression model is telling us and very useful for interpreting and visualizing model estimates. For example, in logistic regression models, the coefficient estimates, when reported as log odds, are not easy to interpret and less informative. Since logistic models are nonlinear (due to the nonlinear link function), in multivariate models with\bmb{?only}\BC{clarify with BMB} interactions, the magnitude of the effect of change in the outcome depends on the values of the predictor of interest and other predictors. Hence, the conclusions one can make about the estimated effects greatly depends on how well we choose the values of the other predictors, i.e., \emph{non-focal} predictors.

Ordinary least squares (OLS) regression provides an easy way to express multivariate relationships as an additive linear combination of many predictors with the outcome. The coefficient estimates from OLS regression models are generally easy to interpret as the expected change in the outcome with a unit change in the continuous predictor or as the expected outcome difference between the levels of the categorical predictor with the baseline category. However, with additional complexities such interaction between predictors, the resulting coefficients risk misinterpretation or interpretation complexity \citep{brambor_understanding_2006, berry_improving_2012}.

Simple OLS regression models are flexible and can easily be generalized, as generalized linear (mixed) models (GL(M)Ms), to examine more complex relationships, including \bmb{nonlinear relationships -- need to clarify}\BC{clarity with BMB} between response and predictors e.g., log transformed response or predictors, interactions between predictors (and via splines for example), and nonlinear transformations via link functions \bmb{due to their flexibility}\BC{clarify why BMB circled it}. This flexibility comes with additional complications, for example, complex multivariate models may risk misinterpretation, and miscalculation of quantities of interest \citep{berry_improving_2012, leeper2017interpreting}. Further, coefficient estimates of GLM models involving nonlinear link functions lack direct interpretation\citep{leeper2017interpreting}, meaning that interpretation of derived quantities from these estimates requires some understanding of the specified model. An alternative is to explore the \emph{effects} of predictors on the predictions of the outcome at various levels of the predictor of interest -- \emph{predictions and effects} \citep{fox2009effect, leeper2017package, lenth2018package}. For example, as a way to plan, public health officials may want to know the effects of household income, wealthindex, etc, on the predicted probability that slum dwellers will have improved water services.

When visually presented, prediction and effects provide a unified and intuitive way of describing relationships from a fitted model, especially complex models involving interaction terms or some kind of transformations on the predictors whose estimates are usually, but not always, harder to interpret. Generating predictions and effects plots together with the associated confidence intervals for regression models has a number of challenges. In particular:
\begin{enumerate}
\item choice of representative values of \emph{focal} predictor and the \emph{reference point} for non-focal predictors especially in multivariate models
\item  propagation of uncertainty -- appropriate choice of \emph{anchor} for computing confidence intervals; and how to incorporate the uncertainty due to non-focal predictors 
\item bias in the expected mean prediction induced by the nonlinear transformation of the response variable (especially in GL(M)Ms)
\end{enumerate}

The most common way of choosing the representative values is taking unique levels of the focal predictor if discrete or taking appropriately sized quantiles if continuous, and then calculating the predictions while holding non-focal (other predictors other than the focal predictor) predictors at their reference point (e.g., means) \citep{hanmer2013behind}. This procedure generates -- \emph{predictions}, \emph{predictor effects} \citep{fox2009effect}, \emph{marginal predictions} \citep{leeper2017package} or \emph{estimated marginal means} \citep{lenth2018package}. In this article, we refer to these quantities as predictions since they should, for example, tell us what we would expect the response to be at a particular value or level of the predictor, for an ``average case''. \bmb{More specifically, we compute the expected outcome by holding non-focal predictors constant (or averaged in some meaningful way) while varying the focal predictor, with the goal that the predictions represents how the model responds to the changes in the focal predictor. On the other hand, effects differ with predictions in the sense that the former is about how we describe the uncertainty around the latter and depends on the anchor. -- see next/unclear}\BC{clarify with BMB}

The most \bmb{commonly used \proglang{R} software packages (\pkg{emmeans} and \pkg{effects})}\BC{not sure the underline} for generating predictions, by default, use the average of non-focal predictors as the reference point. However, there are a number of choices one can make when considering this approach -- for example, in the presence of interaction, averaging the interactions (averaging the product of interacting model variables) versus taking the product of the averages of the interacting model variables (the default for \pkg{emmeans} and \pkg{effects}). For simple models (with no interactions), the two approaches give the same predictions. However, for complex models, we claim that neither of these two approaches is the most appropriate but we show that averaging the interaction closely matches the observed values. To illustrate this, consider models~\ref{eq:simple_inter_higher_no_interaction} and \ref{eq:simple_inter_higher} below, with $x_1$ as the focal predictor:
\bmb{This whole example .... example of nonlinear averaging}\BC{Yes and handling interaction terms?}
%
\begin{align}
y &= \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3x_3 + \epsilon \label{eq:simple_inter_higher_no_interaction}\\
y &= \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3x_3 + \beta_{23}x_2x_3 + \epsilon \label{eq:simple_inter_higher}
\end{align}
%

We first simulate data with the two models, such that $x_{1,2,3} \sim \mathrm{Normal}(0, 1)$, $\beta_0 = 5$, $\beta_1 = -3$, $\beta_2 = 1$, $\beta_3 = 2$ $\beta_{23} = 5$, $\epsilon \sim \mathrm{Normal}(0, \sigma^2)$ and $\sigma^2 = 1$, and then compare the predictions from the \pkg{emmeans}, \pkg{effects} and our proposed alternative (\pkg{varpred}) to the ``true'' predictions and observed average i.e., $\bar{y}$, as shown in Fig~\ref{fig:justify_plots}.

\begin{figure}[!h]
\begin{center}
\includegraphics{justify_preds.inter.pdf}
\end{center}
\caption{{\bf A comparison of \pkg{emmeans}, \pkg{effects} and \pkg{varpred} predictions for $x_1$ on $y$ for models with and without interactions.}
The horizontal blue, green and black lines are the mean predictions, i.e., $\bar{\hat{y}}$, the red ones are the means of the simulated $y$, i.e., $\bar{y}$, while the vertical dotted grey lines are the means of the corresponding focal predictors. The grey points are the binned simulated $y$. The trend lines represent the corresponding $\hat{y}$ at various levels of $x_1$, while holding the other predictors at their average for the three methods and the true predictions based on the simulation parameters. A: In the absence of interaction, the predicted mean, $\bar{\hat{y}}$, closely matches the simulated value in all the three approaches, i.e., $\bar{y} \approx \bar{\hat{y}}$. Similarly, the predictions based on the three approaches closely match the true predictions. B: Even with the simple interaction between non-focal predictors, we start seeing deviation in both predicted mean, $\bar{\hat{y}}$, and overall predictions from the simulated, $\bar{y}$, and true predictions, respectively; in two commonly used packages (\pkg{emmeans} and \pkg{effects}), but not the proposed \pkg{varpred}.}
\label{fig:justify_plots}
\end{figure}

@@

One may be interested in the uncertainties associated with the focal predictor of interest only, excluding other uncertainties due to other non-focal predictors -- effects. However, currently, the two packages do not provide straightforward way do achieve this. To illustrate this, we generate predictions of $x_2$ from model~\ref{eq:simple_inter_higher_no_interaction} together with the associated $95\%$ confidence bands, as shown in Fig~\ref{fig:justify_ci_plots}.

\begin{figure}[!h]
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal predictor, i.e., the model center, corresponds to the effects of the focal predictor from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal predictors. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal predictor.}
\label{fig:justify_ci_plots}
\end{figure}

For \pkg{emmeans} and \pkg{effects}, the confidence bands are much wider because they include uncertainties associated with the intercept and non-focal predictors, but narrower and crosses at the mean of the focal predictor, i.e., \emph{center point or mean-based}, in \pkg{varpred}. In other words, with \pkg{varpred}, we are able to generate effects indicating zero uncertainty at the value of the focal predictor we are more certain about, i.e., mean of the model variable corresponding to the focal predictor (center point or mean-based). For simple models, the point where the confidence bands crosses is the model center; it corresponds to anchor we choose, in this case, the mean of the focal predictor.

When dealing with nonlinear link functions or models with multiple non-focal predictors, the correct predictions are much harder to estimate. One approach is to make predictions on the transformed scale (linear predictor scale), and then back-transform to the original scale. However, the back-transformation may either result in biased predictions or requires some approximation. In particular, bias in expected mean prediction induced by nonlinear transformation of the response variable can lead inaccurate predictions. An alternative to mean-based reference point, i.e., averaging non-focal predictors, is the \emph{whole-sample-based} approach, discussed in detail later, which involves computing the prediction over the population of non-focal predictors and then averaging across the values of the focal predictor \citep{hanmer2013behind}.

@@

\begin{figure}[!h]
\begin{center}
\includegraphics{justify_anchors.combined.pdf}
\end{center}
\caption{{\bf A comparison of min-based and center- (mean-) based anchors.} The center-based effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

We would like to summarize by talking about predictions vs. effects. Predictions are the standard curves; they tell us what we expect to see if we know the value of the focal predictor. The lines inside (effects) are a little different: they are a good way to visualize the effect of the focal predictor. Unlike predictions, where the uncertainty captures the uncertainty in the whole model, the effects plot focuses on uncertainty in the effect of the focal predictor only and depend on the anchor.

@@

To get an intuition of how conditioning on the mean values of non-focal predictors work, suppose we are interested in predictions at the levels of a particular predictor, i.e., focal, $x_f$, from the set of predictors. As a starting point, assume that the model has no interaction terms. The idea is to choose a reference point for the values of non-focal predictors. For example, fixing the values of non-focal predictor(s) at some typical values -- typically determined by averaging (for now) in some meaningful way, for example, arithmetic mean  for continuous and average over the levels of the factors for categorical non-focal predictors. We refer to this as centered or mean-based approach. The most convenient way to achieve this is by constructing centered model matrix, $\bX^\star$, by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.

Consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function (in the case of simple linear model), where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the prediction with respect to the focal predictor in question \citep{fox2009effect}.

An alternative formulation of $\boldetah^\star$ involves expressing the linear predictor as the sum of the focal and non-focal predictors' linear predictors. In particular,

\begin{align}\label{eq:eta_mean}
\eta^\star(x_f, \nset{{\bar{x}^\star}}) &= \hat{\beta}_f x_f + \sum \nset{\hat{\beta}} \nset{{\bar{x}^\star}} \\
\hat{y}_f  &= g^{-1} \left(\eta^\star(x_f, \nset{{\bar{x}^\star}})\right)
\end{align}
where $\nset{{\bar{x}^\star}}$ and $x_f$ are columns of $\bX^\star$ corresponding to the non-focal predictors and the focal predictor, respectively.

@@

Higher order terms such as interactions, splines, polynomials, etc., can be within the focal predictor, within (between) non-focal predictor(s) or between the focal and non-focal predictor(s). To distinguish the three, suppose the model which describes the hypothetical simulated household size simulation based on a number of socio-demographic factors such as age and wealthindex is

@@

In the first case, with \code{Age} as the focal predictor, with cubic polynomial interaction terms (and 3 model variables associated with it). In this case, each of the polynomial terms (model variables) of the focal predictor are evaluated independently across the chosen levels of the focal predictor, $Age_i$. Specifically, the higher order terms are treated as additional columns of the model matrix evaluated with the same values of the focal predictor, while non-focal predictors are fixed at their reference point as discussed in the previous section. In this case, the predictions associated with \code{Age} on the linear predictor scale become

@@

In the second case, with $Wealthindex$ as the focal predictor, the higher order terms in non-focal predictor, \code{Age}, are treated simply as additional columns in the variable space (model variables) and an appropriate choice of reference point applies just like in the models without higher order interactions. For instance, in mean-based approach, we average all non-focal model variables. Thus

\begin{align*}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age\mathop{\vphantom{^2}}}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age\mathop{\vphantom{^2}}}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3}\\
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i
\end{align*}

Lastly, consider model shown in Equation~\ref{eq:simple_inter_higher}, previously used to demonstrate how to handle interactions in non-focal predictors. Here, we use the same model to demonstrate how to handle interactions between the focal predictor, $x_2$, and non-focal predictor, $x_3$. In this case, the non-interacting predictors are treated as non-focal predictors, with appropriate reference points. However, for the interacting predictors, we first choose representative values for the focal predictor and some particular values of the interacting predictor similar to the way we choose values for the focal predictor or predetermined. In our example, suppose we pick $i$ and $j$ unique values of the focal predictor, $x_2$ and interacting predictor $x_3$, respectively. The prediction on linear predictor scale is given by

@@

We describe the uncertainty around the estimates using confidence intervals (CI). In principle, every value of focal predictor has a different CI. The conventional way to compute variances for predictions is $\boldsymbol\sigma^2 = \textrm{Diag}(\bX^\star \boldsymbol{\Sigma} \bX^{\star\top})$ \citep{lenth2018package, fox2009effect}, so that the confidence intervals are $\boldeta \pm q\boldsymbol\sigma$, where $q$ is an appropriate quantile of Normal or t distribution. This generates prediction CIs and incorporates all the uncertainties -- including the uncertainty due to the intercept and non-focal predictors.  But what if we are only interested in the uncertainty as a result of the focal predictor, so that the confidence intervals are $\boldeta \pm q \boldsymbol \sigma_f$, i.e., effects?

Commonly used \proglang{R} packages for constructing predictions do not exclude the uncertainties resulting from non focal predictors when computing the CIs. Currently a non-trivial way to exclude uncertainties associated with non-focal predictors in some of these packages is to provide a user defined variance-covariance matrix with the covariances of non-focal terms set to $0$. We refer to this procedure as \emph{zeroing-out} variance-covariance matrix; and only works when the input predictors are \emph{centered} prior to model fitting, in case of numerical predictors, and more complicated when the predictors are categorical. We describe our proposed method which is based on the model center and does not require input predictors to be centered prior to model fitting.

@@

Let $\bxo$ be a model matrix containing appropriately chosen values of focal predictor and centered non-focal model variables, and let $\ba$ be the anchor matrix, with same dimension and entries in all non-focal model variables as $\bxo$. Let $\baf$ be the column of $\ba$ corresponding to focal predictor model variable(s) defined in $\bxo$. Any appropriate values can be chosen for $\baf$ but for model center (center-anchored), we use $\bafc$ which is the mean of the focal predictor model variable(s). Thus the variances for the predictions we defined in the previous section is

%
\begin{align}\label{eq:centered_variance}
\boldsymbol\sigma_f^2 = \textrm{Diag}((\bxo - \bafc) \boldsymbol{\Sigma} (\bxo - \bafc)^\top)
\end{align}
%
We can see that $\forall ~\bxof=\baf$, $\bxo - \baf = \boldsymbol 0$, hence $\boldsymbol\sigma_f^2 = \boldsymbol{0}$. Similarly, for all values of $\bxof$ close to anchor point $\bafc$ the term $(\bxo - \bafc)$ and $\boldsymbol\sigma_f^2$ goes to $\boldsymbol 0$ and $\boldsymbol\sigma_f^2 = \boldsymbol 0$ if $\bxof=\bafc$. This means that $\boldsymbol\sigma_f^2$ close to the anchor point are smaller than those away from the anchor point; and results to confidence intervals which are narrower (isolated) around the anchor or crosses at the anchor point for simple models.

The computation of $\bxo - \bafc$ impacts only on the intercepts and non-focal model variables, i.e., the slopes and variance corresponding to the focal predictors are not affected. This means that we can still generate effects without necessarily centering the predictors prior to model fitting.

@@

It is usually important to report the estimates that reflect the expected values of the untransformed response. This is not a major issue in simple linear models. However, when dealing with nonlinear link functions,  generated predictions may not reflect the observed response due to the bias in the expected mean induced by the nonlinear transformation of the response variable. In such cases, bias correction is needed when back-transforming the predictions to the original scales. The common approach for bias-adjustment is second-order Taylor approximation \citep{lenth2018package, duursma2003bias}. Another potential source of bias comes from additional non-focal predictors, especially in models with nonlinear link functions. Here, we describe and implement a different approach; whole-sample-based approach for bias correction.

@@

One way to predict is to condition on values of the focal predictor and make predictions for all observations (members of the population) \citep{hanmer2013behind}. The nonlinear transformation involved in these computations is always \emph{one-dimensional}; all of the multivariate computations required are at the stage of collapsing the multidimensional set of predictors for some subset of the population to a one-dimensional distribution of $\eta^\star(x_f, \nset{x})$, which is a function of focal predictor and the observed values of non-focal predictors, as opposed to the definition in Equation~\ref{eq:eta_mean}. More specifically:
